---
title: "Pivot Lobsters"
author: "Jerry Ngai"
date: "1/27/2020"
output: html_document
---

Attaching our packages and reading in data.

```{r, warning=FALSE, message=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) # 
library(here) # 
library(skimr) # install.packages("skimr")
library(kableExtra) # install.packages("kableExtra")

# read in lobster data
lobsters <- read_xlsx(here("data", "lobsters2.xlsx"), 
                      skip = 4)



```


Viewing our data with skimr

```{r, warning=FALSE}
# these are equivalent
skimr::skim(lobsters) # explicitly stated from skimr
skim(lobsters) # not explicit

```

## Pivot Tables
In R, Pivot tables are: `group_by() %>% summarize ()`
The `n()` function will count observations.
```{r}

# group by year
lobsters %>% 
  group_by(year) %>%
  summarize(count_by_year = n())

# group by year and site 
lobsters %>% 
  group_by(year, site) %>%
  summarize(count_by_yearsite = n())

# group by site and year 
lobsters %>% 
  group_by(site, year) %>%
  summarize(count_by_siteyear = n())
```

Summarize by multiple variables and assign as an object
```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>%
  summarize(count_by_siteyear = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))

```

Table formatting with `kableExtra::kable()`
```{r}

siteyear_summary %>% 
  kable()
```

There are `r nrow(lobsters)` total lobsters included in this report.

## Activity

1. build from our analyses and calculate the medium lobster size for each site and year
```{r}
siteyear_summary2 <- lobsters %>% 
  group_by(site, year) %>%
  summarize(count_by_siteyear = n(),
            median_size_mm = median(size_mm, na.rm = TRUE))

siteyear_summary2 %>% 
  kable()
```

2. crete a ggplot and save. For example:
  
```{r}
ggplot(data = siteyear_summary2, aes(x=year, y=median_size_mm, color = site)) +
  geom_line()

```

3. sync this to github