---
title: "Creating graphs with ggplot2"
author: "Jerry Ngai"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.

We will install these packages:

- tidyverse: suite of packages for wrangling & viz (ggplot)
- readxl: to read in excel files (in may ways)
- here: to help with filepaths

```{r setup, message=FALSE, warning=FALSE}
library (tidyverse) # install.packages("tidyverse")
library (readxl) #install.packages("readxl")
library (here) #install.packages("here")

```

# California National Parks Data

```{r read-data}

ca_np <- read_csv(here("data/ca_np.csv"))
# can also be written:
# ca_np <- read_csv("data/ca_np.csv")
# ca_np <- read_csv(here("data", "ca_np.csv"))

ci_np <- read_excel(here("data", "ci_np.xlsx"))

```

## Make my own first graph (Plot of Channel Islands data)

We're going to make a line plot of year (x-variable) vs visitors (y-variable) in ggplot:

```{r plot-data}

ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_line()

ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point()

# Update the geom_ to geom_point(), geom_col(), geom_area()

```

Make a base layer for my CI plots:

```{r}

gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point()

```

## Now lets start customizing!  (based on the Grammar of Graphics)

- Line / point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparaency: alpha

```{r}

gg_base + 
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
  
```

Activity: create a ggplot scatterplot (geom_point) that has triangle symols (hint: ?pch), of a non-default size (size = #), and that are an update color.


```{r}

gg_base + 
  geom_point(
    color = "orange4",
    size = 20,
    pch = 2
  )
  
```

## Update graph aesthetics based on variables

```{r}

gg_base + 
  geom_point(
    aes(size = visitors)
  )

```

# Add an argument to update the color based on the year value

```{r}

gg_base + 
  geom_point(
    aes(size = visitors, 
          color = year),
    pch = 10
  )

```



