---
title: "Tidying Session"
author: "Jerry Ngai"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attaching packages

library(tidyverse) # install.packages("tidyverse")
library(janitor) # install.packages("janitor")
library(here) # install.packages("here")
library(readxl) # install.packages("readxl")

```

## Read files

### Read this file: inverts.xlsx

```{r}
inverts <- read_excel(here("data", "inverts.xlsx"))

```

### `pivot_longer()`
Use single colon (:) to indicate a series of sequential columns
```{r}

inverts_long <- pivot_longer(data = inverts,
                             cols = '2016':'2018',
                             names_to = "year", 
                             values_to = "sp_count")
```

What if we want to get it back into wide format?

```{r}
inverts_wide <- inverts_long %>%
  pivot_wider(names_from = common_name,
              values_from = sp_count)

```

### `janitor::clean_names()`

```{r}
inverts_wide <- inverts_wide %>%
  clean_names()
```

### `tidyr::unite()` to combine information from columns

Example: merge the site code and year into a single column called "site_year"
```{r}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year",
        c(site, year),
        sep = "_")
```

### Activity:

Create a new object called 'inverts_mo_yr' starting from `inverts_long` that unites the month and year columns into a SINGLE column named "mo_yr", using a slash "/" as the separator...then try some other fun separator.

```{r}
inverts_unite2 <- inverts_long %>% 
  unite(col = "mo_yr",
        c(month, year),
        sep = "/")

inverts_unite3 <- inverts_long %>% 
  unite (col = "site_mo_year",
         c(site, month, year),
         sep = "^.^")
```


### `tidyr::separate()`

```{r}
inverts_sep <- inverts_unite %>%
  separate(col = site_year,
           into = c("my_site", "my_year"))
```

### `stringr` + `mutate()`

Example: using built-in `trees` dataset with mutate() to add 5 to all the tree heights

```{r}
trees_new <- trees %>%
  mutate(height_update = Height + 5)
```

### `str_replace`

```{r}
ca_abbr <- inverts %>%
  mutate(common_name =
           str_replace(common_name,
                       pattern = "california",
                       replacement = "CA"))

```

Example: In the common_name column, you want to replace the word "spiny" with the word "awesome"
Store as `awesome_lobsters`

```{r}
awesome_lobsters <- inverts %>%
  mutate(common_name =
           str_replace(common_name,
                       pattern = "spiny",
                       replacement = "AWESOME"))

```